---
title: "Decision Trees and Random Forests"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

***Problem Statement:***
*This project involves building decision tree and random forest models to answer a number of questions. We will use the Carseats dataset that is part of the ISLR package.*

**Loading Required Libraries**
```{r}

library(ISLR)

library(dplyr)

library(glmnet)

library(caret)

library(rpart)

library(rpart.plot)

library(rattle)
```

**Selecting required attributes**
```{r}

Carseats_Filtered <- Carseats %>% select("Sales", "Price", 
"Advertising","Population","Age","Income","Education")

```


**Building a decision tree regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education").**

```{r}

model_1<- rpart(Sales~.,data=Carseats_Filtered,method = 'anova')

```

```{r}
fancyRpartPlot(model_1)

```
*Price attribute is used at the top of the tree (root node) for the splitting.*


**Considering the following input: Sales=9,Price=6.54,Population=124,Advertising=0,Age=76,Income= 110,Education=10.Estimating Sales for this record using the decision tree model**

```{r}

prediction_data = data.frame(Price=6.54 ,Population=124,Advertising=0,Age=76
,Income= 110, Education= 10)


prediction<- predict(model_1,prediction_data)

prediction

```



*Predicted sales value for this record is 9.58625.*

**Using the caret function to train a random forest (method=’rf’) for the same dataset.**

```{r}

set.seed(123)
model_2 <- train(Sales~.,
               data= Carseats_Filtered,
               method = "rf")

# Print the results
model_2

```

*Best results are obtained when mtry value is set to be 2.*

**Customizing the search grid by checking the model’s performance for mtry values of 2, 3 and 5 using 3 repeats of 5-fold cross validation.**
```{r}

set.seed(123)

#Cross-Validation
control <- trainControl(method = "repeatedcv",
                     repeats = 3,
                     number = 5)

#Defining search grid with mtry values of 2,3,5
search_grid <- expand.grid(mtry = c(2, 3, 5))


# Training the model using the search grid and cross-validation
model <- train(Sales~., Carseats_Filtered,method = "rf",tuneGrid = search_grid,trControl = control)
print(model)

```
