---
title: "Online Retail Analytics-Data Transformations"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```
***Purpose of this Project:***

*The main purpose of this project is to apply Data Transformation on 'Online Retail Dataset' to analyze the role of Descriptive Statistics in EDA.*

```{r}

#Loading required library:
library(dplyr)

#Reading the Dataset:
Online_Retail<-read.csv("./Online_Retail.csv")
```
**1. Breakdown of the number of transactions by countries in both percentage and count, and showing the countries accounting for more than 1% of the total transactions:**
```{r}

trans_countries<-Online_Retail %>% group_by(Country) %>% summarise(cnt = n()) %>% mutate(perc =round((cnt/sum(cnt))*100,5)) %>% filter(perc>1)
head(trans_countries) 

```
**2. Creating a new variable TransactionValue and adding it to the dataframe:**
```{r}
TransactionValue<-Online_Retail$Quantity*Online_Retail$UnitPrice

#creating a dataframe and adding TransactionValue to it

Online_Retail_new<-data.frame(InvoiceNo=Online_Retail$InvoiceNo,StockCode= Online_Retail$StockCode,Description=Online_Retail$Description,Quantity=Online_Retail$Quantity, InvoiceDate=Online_Retail$InvoiceDate,UnitPrice=Online_Retail$UnitPrice, CustomerID=Online_Retail$CustomerID,Country=Online_Retail$Country,  TransactionValue=TransactionValue)


```

**3. Showing the breakdown of transaction values by countries in total sum of transaction value. Displaying countries with total transaction exceeding 13000**
```{r}

Trans_value_countries<- Online_Retail_new %>% group_by(Country) %>% summarise(sum_TransactionValue = sum(TransactionValue)) %>% filter(sum_TransactionValue>130000)

head(Trans_value_countries)

```
**4. Optional question**
```{r}
Temp=strptime(Online_Retail_new$InvoiceDate,format='%m/%d/%Y %H:%M',tz='GMT')
head(Temp)

#New_Invoice_Date
Online_Retail_new$New_Invoice_Date<- as.Date(Temp)

Online_Retail_new$New_Invoice_Date[20000]- Online_Retail_new$New_Invoice_Date[10]

#Invoice_Week

Online_Retail_new$Invoice_Day_Week= weekdays(Online_Retail_new$New_Invoice_Date)

#Invoice_Hour

Online_Retail_new$New_Invoice_Hour = as.numeric(format(Temp, "%H"))

#Invoice_month

Online_Retail_new$New_Invoice_Month = as.numeric(format(Temp, "%m"))

```

**4(a). Percentage of transactions (by numbers) by days of the week**

```{r}
perc_transc<- Online_Retail_new %>% group_by(Invoice_Day_Week) %>% summarise(count=n()) %>% mutate(perc =round((count/sum(count))*100,5))

head(perc_transc)

```
**4(b). Percentage of transactions (by transaction volume) by days of the week**
```{r}

perc_trans_week<- Online_Retail_new %>% group_by(Invoice_Day_Week) %>% summarise(Total=sum(TransactionValue)) %>% mutate(percentage = (Total/sum(Total))*100)
head(perc_trans_week)

```
**4(c). Percentage of transactions (by transaction volume) by month of the year**

```{r}
perc_trans_month<- Online_Retail_new %>% group_by(New_Invoice_Month) %>% summarise(Total=sum(TransactionValue)) %>% mutate(percentage = (Total/sum(Total))*100)
head(perc_trans_month)

```

**4(d). The date with the highest number of transactions from Australia**

```{r}
date_trans<- Online_Retail_new %>% filter(Country == 'Australia') %>% group_by(New_Invoice_Date) %>% summarise(count=n()) %>% arrange(desc(count))

head(date_trans)

#Australia recorded highest number of transactions on 2011-06-15
```

**5. Histogram of transaction Values from Germany**
```{r}

histogram<- Online_Retail_new %>% filter(Country == 'Germany') 
numeric_TV<- as.integer(histogram$TransactionValue)
hist(numeric_TV)

```

**6. Identifying the customer with highest number of transactions and finding the most valuable customer**
```{r}

cust_count<-Online_Retail_new %>% group_by(CustomerID) %>% summarise(cntt = n()) %>% arrange(desc(cntt))
head(cust_count)

#Customer 17841 has the highest number of transactions.

```

```{r}
cust_sum<-Online_Retail_new %>% group_by(CustomerID) %>% summarise(sum_cnt =sum(TransactionValue)) %>% arrange(desc(sum_cnt))

head(cust_sum)

#Customer 14646 is the most valuable

```

**7. Percentage of missing values for each variable in the dataset**
```{r}
missing_values<- (colMeans(is.na(Online_Retail_new))*100)

missing_values
```

**8. Number of transactions with missing CustomerID records by countries**
```{r}

missing<-Online_Retail_new %>% filter(is.na(CustomerID)) %>% group_by(Country) %>% summarise(count_by_countries = n())

View(missing)

```

**9. On average, how often the costumers comeback to the website for their next shopping?**

```{r}

difference_days<-Online_Retail_new %>% select(CustomerID,New_Invoice_Date) %>% group_by(CustomerID) %>% distinct(New_Invoice_Date) %>% arrange(desc(CustomerID)) %>% mutate(days = New_Invoice_Date-lag(New_Invoice_Date)) %>% filter(!is.na(days)) 
View(difference_days)
mean(difference_days$days)

# On an average,customers come back after 38 days to the website for their next shopping.
```

**10. Return rate for the French customers**

```{r}
cancelled_customers <- Online_Retail_new %>% filter(Country=='France',Quantity<0) %>% summarise(count = n())
Total_customers<-  Online_Retail_new %>% filter(Country=='France') %>% count()

return_rate_french_cust=((cancelled_customers/Total_customers)*100)

head(return_rate_french_cust)

```

**11. Product that has generated the highest Revenue for the retailer**
```{r}

item_sum<-Online_Retail_new %>% group_by(Description) %>% summarise(sum_cnt = sum(TransactionValue)) %>% arrange(desc(sum_cnt))
head(item_sum)

#DOTCOM POSTAGE generates highest revenue for the retailer
```

**12. Unique customers in the dataset**
```{r}

unique_cust<- Online_Retail_new %>% distinct(CustomerID) %>% summarise(ncount = n())
head(unique_cust)
 #There are 4373 unique customers in the dataset
```
