---
title: "Support Vector Machines and Neural Networks"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

***Purpose of this Project:***

*This project involves building SVM and neural network regression models to answer a number of questions. We will use the Carseats dataset that is part of the ISLR package.*

**Loading required libraries:**
```{r}
library(ISLR)
library(dplyr)
library(glmnet)
library(caret)
```

**Selecting required features:**
```{r}
Carseats_Filtered <- Carseats %>% select("Sales", "Price", 
"Advertising","Population","Age","Income","Education")


```


**Building a linear SVM regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education").:**
```{r}

set.seed(1203)
SVM_Model<- train(Sales~.,data=Carseats_Filtered,method="svmLinear",preProcess=c("center","scale"),tuneLength = 10)
SVM_Model

```
*R square is 34.38%*

**Customizing the search grid by checking the model’s performance for C parameter of 0.1,.5,1 and 10 using 2 repeats of 5-fold cross validation. **

```{r}

set.seed(1203)

grid = expand.grid(C= c(0.1,0.5,1,10))

trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

SVM_Model2 <- train(Sales ~., data = Carseats_Filtered, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"), tuneGrid = grid,
tuneLength = 10)

SVM_Model2
```
**Training a neural network model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). **
```{r}
Normalization <- preProcess(Carseats_Filtered[,2:7],method = c("center","scale"))
Norm_data<-predict(Normalization,Carseats_Filtered)

NNET_Model<- train(Sales~.,data=Norm_data,method="nnet",linout=TRUE, trace = FALSE)
NNET_Model
```
*R-square value with best hyperparameters( size=1) is 33.43.*

**Consider the following input: Sales=9,Price=6.54,Population=124,Advertising=0,Age=76,Income= 110,Education=10. Estimating Sales for this record using the above neuralnet model? **
```{r}
Input <- data.frame(Price = 6.54, Population = 124, Advertising = 0, Age = 76, Income = 110, Education = 10)

Prediction<- predict(NNET_Model,Input )
Prediction
```

*Estimated sales: 4.976281*
